Distributional Trees and Forests

### 2017-06-22 pkg/disttree/inst/useR2017/slides.Rnw
modified plots
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1476 255f4249-2263-4417-b1ea-8498779e27a7

coltree <- hcl(c(128,100,30))
colforest <- hcl(c(10,100,30))
colgamlss <- hcl(c(260,100,30))

@@

  # disttree
  if(dt.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "disttree", xaxt='n', yaxt='n', xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.dt, type = "l", col = coltree)
    lines(x = sp$x, y = sp$fitted.mu.dt + sp$fitted.sigma.dt, type = "l", col = 'greenyellow')
    lines(x = sp$x, y = sp$fitted.mu.dt - sp$fitted.sigma.dt, type = "l", col = 'greenyellow')
    legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
                        TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')), 
           col = c('black','grey',coltree, 'greenyellow'), lty = 1, bty = 'n')
  }

  # distforest
  if(df.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "distforest", xaxt='n', yaxt='n', xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.df, type = "l", col = colforest)
    lines(x = sp$x, y = sp$fitted.mu.df + sp$fitted.sigma.df, type = "l", col = 'orange')
    lines(x = sp$x, y = sp$fitted.mu.df - sp$fitted.sigma.df, type = "l", col = 'orange')
    legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
                        TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')), 
           col = c('black','grey',colforest, 'orange'), lty = 1, bty = 'n')
  }

  # gamlss
  if(g.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey",  main = "gamlss", xaxt='n', yaxt='n', xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.g, type = "l", col = colgamlss)
    lines(x = sp$x, y = sp$fitted.mu.g + sp$fitted.sigma.g, type = "l", col = 'lightblue')
    lines(x = sp$x, y = sp$fitted.mu.g - sp$fitted.sigma.g, type = "l", col = 'lightblue')
    legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
                        TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')),  
           col = c('black','grey',colgamlss, 'lightblue'), lty = 1, bty = 'n')
  }

  # disttree vs distforest vs gamlss
  if(all.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "disttree vs distforest vs gamlss", xaxt='n', yaxt='n', xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    lines(x = sp$x, y = sp$fitted.mu.dt, type = "l", col = coltree)
    lines(x = sp$x, y = sp$fitted.mu.df, type = "l", col = colforest)
    lines(x = sp$x, y = sp$fitted.mu.g, type = "l", col = colgamlss)
    legend('topleft', c(TeX('$\\mu$'), TeX('$\\hat{\\mu}$ disttree'),
                        TeX('$\\hat{\\mu}$ distforest'), TeX('$\\hat{\\mu}$ gamlss')), 
           col = c('black',coltree,colforest,colgamlss), lty = 1, bty = 'n')
  }
}

@@

\setkeys{Gin}{width=0.74\linewidth}
<<plottree_dgp, fig=TRUE, echo=FALSE>>=
pal <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
plot(fig, inner_panel = node_inner_ext,
     tp_args = list(FUN = identity, width = 18, fill = pal[c(1, 3)]), 
     ip_args = list(fill = pal[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1.4)
@
\end{center}
\end{frame}

<<tree, echo=FALSE, results=hide>>=
set.seed(7)
nobs <- 500
x <- runif(nobs, 0, 1)
mu <- sigma <- ytrue <- numeric(length = nobs)
for(i in 1:nobs) sigma[i] <- if(x[i]<=0.4) 1 else 3
for(i in 1:nobs) mu[i] <- if(x[i]<= 0.4|| x[i]>0.8) 4 else 12
y <- rnorm(nobs, mean = mu, sd = sigma)
#y <- rcnorm(nobs, mean = mu, sd = sigma, left = 0)
ytrue <- mu
data <- data.frame(cbind(y,x, ytrue))
tree <- disttree(y ~ x, data = data, family = NO(), type.tree = "mob")
#tree <- disttree(y ~ x, data = data, family = dist_list_cens_normal)

#FUN <- function(x) {
#                cf <- x$coefficients
#                cf <- matrix(cf, ncol = 1, dimames = list(c(expression(mu), expression(sigma)), ""))
#                c(sprintf("n = %s", x$nobs), "Estimated parameters:", 
#                  strwrap(capture.output(print(cf, digits = 4L))[-1L]))
#}

#FUN <- function(x) {
#                cf <- x$coefficients
#                cf <- matrix(cf, ncol = 1, dimnames = list(names(cf), 
#                  ""))
#                c(sprintf("n = %s", x$nobs), "Estimated parameters:", 
#                  strwrap(capture.output(print(cf, digits = 4L))[-1L]))
#            }
#FUN <- function (x) 
#{
#  cf <- x$coefficients
#  cf <- matrix(cf, ncol = 1, dimnames = list(names(cf), ""))
#  c(sprintf("n = %s", x$nobs), "Estimated parameters:", c(c(expression(mu), round(cf[1],3)), 
#                                                          c(expression(sigma), round(cf[2],3))))
#}
@

\begin{frame}[fragile]
\frametitle{Fitting a Tree}
\begin{center}
\vspace*{-0.5cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.74\linewidth}
<<plottree_estpar, fig=TRUE, echo=FALSE>>=
pal <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
plot(tree, drop = TRUE, tnex = 1.4, #FUN = FUN,
     tp_args = list(fill = pal[c(1, 3)]), 
     ip_args = list(fill = pal[c(2, 3)]))

@
\end{center}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Fitting a Tree}
\begin{center}
\vspace*{-0.5cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.74\linewidth}
<<plottree_box, fig=TRUE, echo=FALSE>>=
plot(as.constparty(tree), tnex = 1.4,
     tp_args = list(fill = pal[c(1, 3)]), 
     ip_args = list(fill = pal[c(2, 3)]))
@
\end{center}
\end{frame}

@@

  if(type == "exp"){
    #dev.off() 
    rmse <- cbind(simlist$av.rmse.exp.true.dt, simlist$av.rmse.exp.true.df, simlist$av.rmse.exp.true.g,
                  simlist$av.rmse.exp.obs.dt, simlist$av.rmse.exp.obs.df, simlist$av.rmse.exp.obs.g)
    colnames(rmse) <- c("dt.true", "df.true", "g.true", "dt.obs", "df.obs", "g.obs")
    ylim <- c(min(rmse), max(rmse))
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(x = simlist$x.axis, y = rmse[,"dt.true"], type = "l", col = coltree, ylim = ylim,
         xlab = "", ylab = "", main = "disttree vs distforest vs gamlss", xaxt='n', yaxt='n')
    lines(x = simlist$x.axis, y = rmse[,"dt.obs"], type = "l", lty = 2, col = coltree)
    lines(x = simlist$x.axis, y = rmse[,"df.true"], type = "l", col = colforest)
    lines(x = simlist$x.axis, y = rmse[,"df.obs"], type = "l", lty = 2, col = colforest)
    lines(x = simlist$x.axis, y = rmse[,"g.true"], type = "l", col = colgamlss)
    lines(x = simlist$x.axis, y = rmse[,"g.obs"], type = "l", lty = 2, col = colgamlss)
    #legend('left', c("disttree", "distforest", "gamlss"), col = c(coltree, colforest, colgamlss), lty = 1)
    mtext(text= "smooth", side = 1, line = 1, adj = 0)
    mtext(text= "steep", side = 1, line = 1, adj = 1)
    mtext(text= TeX('$\\kappa$'), side = 1, line = 1)
    mtext(text= "RMSE", side = 2, line = 1)
    mtext(text= "disttree", side = 2, col = coltree, las = 1, line = 0.2, padj = 6.5)
    mtext(text= "distforest", side = 2, col = colforest, las = 1, line = 0.2, padj = 13)
    mtext(text= "gamlss", side = 2, col = colgamlss, las = 1, line = 0.3, padj = 16)

@@

    #add arrows below
    arrows(0.45,0.09,0.3,0.09,lwd=1) ## add arrow
    arrows(0.6,0.09,0.75,0.09,lwd=1) ## add arrow

@@

  if(type == "par"){
    rmse <- cbind(simlist$av.rmse.mu.dt, simlist$av.rmse.mu.df, simlist$av.rmse.mu.g, 
                  simlist$av.rmse.sigma.dt, simlist$av.rmse.sigma.df, simlist$av.rmse.sigma.g)
    colnames(rmse) <- c("mu.dt", "mu.df", "mu.g", "sigma.dt", "sigma.df", "sigma.g")
    ylim <- c(min(rmse), max(rmse))
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(x = simlist$x.axis, y = rmse[,"mu.dt"], type = "l", col = coltree, ylim = ylim,
         xlab = "", ylab = "", main = "disttree vs distforest vs gamlss", xaxt='n', yaxt='n')
    lines(x = simlist$x.axis, y = rmse[,"sigma.dt"], type = "l", lty = 2, col = coltree)
    lines(x = simlist$x.axis, y = rmse[,"mu.df"], type = "l", col = colforest)
    lines(x = simlist$x.axis, y = rmse[,"sigma.df"], type = "l", lty = 2, col = colforest)
    lines(x = simlist$x.axis, y = rmse[,"mu.g"], type = "l", col = colgamlss)
    lines(x = simlist$x.axis, y = rmse[,"sigma.g"], type = "l", lty = 2, col = colgamlss)
    #legend('topleft', c("disttree", "distforest", "gamlss"), col = c(coltree, colforest,colgamlss), lty = 1)
    mtext(text= "smooth", side = 1, line = 1, adj = 0)
    mtext(text= "steep", side = 1, line = 1, adj = 1)
    mtext(text= TeX('$\\kappa$'), side = 1, line = 1)
    mtext(text= "RMSE", side = 2, line = 1)
    mtext(text= "disttree", side = 2, col = coltree, las = 1, line = 0.2, padj = -4)
    mtext(text= "distforest", side = 2, col = colforest, las = 1, line = 0.2, padj = 8)
    mtext(text= "gamlss", side = 2, col = colgamlss, las = 1, line = 0.3, padj = 14)

@@

# plot loglikelihood
plot_ll <- function(simlist){
  ll <- cbind(simlist$av.loglik.dt, simlist$av.loglik.df, simlist$av.loglik.g)
  colnames(ll) <- c("dt", "df","g")
  ylim <- c(min(ll), max(ll))
  par(mar=c(5.1,4.1,4.1,3.1))
  plot(x = simlist$x.axis, y = ll[,"dt"], type = "l", col = coltree, ylim = ylim,
       xlab = "", ylab = "", main = "disttree vs distforest vs gamlss", xaxt='n', yaxt='n')
  lines(x = simlist$x.axis, y = ll[,"df"], type = "l", col = colforest)
  lines(x = simlist$x.axis, y = ll[,"g"], type = "l", col = colgamlss)
  #legend('topright', c("disttree", "distforest", "gamlss"), col = c(coltree, colforest,colgamlss), lty = 1)
  mtext(text= "smooth", side = 1, line = 1, adj = 0)
  mtext(text= "steep", side = 1, line = 1, adj = 1)
  mtext(text= TeX('$\\kappa$'), side = 1, line = 1)
  mtext(text= "Log-Likelihood", side = 2, line = 1)
  mtext(text= "disttree", side = 2, col = coltree, las = 1, line = 0.2, padj = 12)
  mtext(text= "distforest", side = 2, col = colforest, las = 1, line = 0.2, padj = -8.5)
  mtext(text= "gamlss", side = 2, col = colgamlss, las = 1, line = 0.3, padj = -16)

>>>>

pal <- hcl(c(10, 128, 260), 100, 50)
names(pal) <- c("forest", "tree", "gamlss")

@@

  # disttree
  if(dt.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "disttree", xaxt='n', yaxt='n', xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.dt, type = "l", col = pal["tree"], lwd = 2)
    lines(x = sp$x, y = sp$fitted.mu.dt + sp$fitted.sigma.dt, type = "l", col = 'greenyellow')
    lines(x = sp$x, y = sp$fitted.mu.dt - sp$fitted.sigma.dt, type = "l", col = 'greenyellow')
    legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
                        TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')), 
           col = c('black','grey',pal["tree"], 'greenyellow'), lty = 1, bty = 'n')
  }

  # distforest
  if(df.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "distforest", xaxt='n', yaxt='n', xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.df, type = "l", col = pal["forest"], lwd = 2)
    lines(x = sp$x, y = sp$fitted.mu.df + sp$fitted.sigma.df, type = "l", col = 'orange')
    lines(x = sp$x, y = sp$fitted.mu.df - sp$fitted.sigma.df, type = "l", col = 'orange')
    legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
                        TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')), 
           col = c('black','grey',pal["forest"], 'orange'), lty = 1, bty = 'n')
  }

  # gamlss
  if(g.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey",  main = "gamlss", xaxt='n', yaxt='n', xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.g, type = "l", col = pal["gamlss"], lwd = 1.5)
    lines(x = sp$x, y = sp$fitted.mu.g + sp$fitted.sigma.g, type = "l", col = 'lightblue')
    lines(x = sp$x, y = sp$fitted.mu.g - sp$fitted.sigma.g, type = "l", col = 'lightblue')
    legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
                        TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')),  
           col = c('black','grey',pal["gamlss"], 'lightblue'), lty = 1, bty = 'n')
  }

  # disttree vs distforest vs gamlss
  if(all.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "disttree vs distforest vs gamlss", xaxt='n', yaxt='n', xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    lines(x = sp$x, y = sp$fitted.mu.dt, type = "l", col = pal["tree"], lwd = 2)
    lines(x = sp$x, y = sp$fitted.mu.df, type = "l", col = pal["forest"], lwd = 2)
    lines(x = sp$x, y = sp$fitted.mu.g, type = "l", col = pal["gamlss"], lwd = 2)
    legend('topleft', c(TeX('$\\mu$'), TeX('$\\hat{\\mu}$ disttree'),
                        TeX('$\\hat{\\mu}$ distforest'), TeX('$\\hat{\\mu}$ gamlss')), 
           col = c('black',pal["tree"],pal["forest"],pal["gamlss"]), lty = 1, bty = 'n')
  }
}

@@

\setkeys{Gin}{width=0.74\linewidth}
<<plottree_dgp, fig=TRUE, echo=FALSE>>=
pal2 <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
plot(fig, inner_panel = node_inner_ext,
     tp_args = list(FUN = identity, width = 18, fill = pal2[c(1, 3)]), 
     ip_args = list(fill = pal2[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1.4)
@
\end{center}
\end{frame}

<<tree, echo=FALSE, results=hide>>=
set.seed(7)
nobs <- 500
x <- runif(nobs, 0, 1)
mu <- sigma <- ytrue <- numeric(length = nobs)
for(i in 1:nobs) sigma[i] <- if(x[i]<=0.4) 1 else 3
for(i in 1:nobs) mu[i] <- if(x[i]<= 0.4|| x[i]>0.8) 4 else 12
y <- rnorm(nobs, mean = mu, sd = sigma)
#y <- rcnorm(nobs, mean = mu, sd = sigma, left = 0)
ytrue <- mu
data <- data.frame(cbind(y,x, ytrue))
tree <- disttree(y ~ x, data = data, family = NO(), type.tree = "mob")
#tree <- disttree(y ~ x, data = data, family = dist_list_cens_normal)
@

\begin{frame}[fragile]
\frametitle{Fitting a Tree}
\begin{center}
\vspace*{-0.5cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.74\linewidth}
<<plottree_estpar, fig=TRUE, echo=FALSE>>=
# function for output in terminal panels
FUN <- function (x) 
{
  cf <- x$coefficients
  cf <- matrix(cf, ncol = 1, dimnames = list(names(cf), ""))
  c(sprintf("n = %s", x$nobs), "Estimated parameters:", parse(text = paste("mu == ", round(cf[1],3))), 
                                                        parse(text = paste("sigma == ", round(cf[2],3))))
}

# alternative to using FUN:
# change info slot in each terminal node and use FUN = identity
if(FALSE){
info2 <- tree$node$kids[[1]]$info 
info4 <- tree$node$kids[[2]]$kids[[1]]$info 
info5 <- tree$node$kids[[2]]$kids[[2]]$info

cf <- info2$coefficients
info2 <- c(sprintf("n = %s", info2$nobs), "Estimated parameters:", parse(text = paste("mu == ", round(cf[1],3))), 
                                                          parse(text = paste("sigma == ", round(cf[2],3))))

cf <- info4$coefficients
info4 <- c(sprintf("n = %s", info4$nobs), "Estimated parameters:", parse(text = paste("mu == ", round(cf[1],3))), 
                                                          parse(text = paste("sigma == ", round(cf[2],3))))

cf <- info5$coefficients
info5 <- c(sprintf("n = %s", info5$nobs), "Estimated parameters:", parse(text = paste("sigma == ", round(cf[2],3))), 
                                                          parse(text = paste("mu == ", round(cf[1],3))))

tree$node$kids[[1]]$info <- info2
tree$node$kids[[2]]$kids[[1]]$info <- info4 
tree$node$kids[[2]]$kids[[2]]$info <- info5

class(tree) <- "party"
}

pal2 <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))

## plot version using FUN and tree of class 'disttree'
plot(tree, drop = TRUE, tnex = 1.4, FUN = FUN,
     tp_args = list(fill = pal2[c(1, 3)], width = 18), 
     ip_args = list(fill = pal2[c(2, 3)]))

## plot version using replaced info slots and tree of class 'party'
if(FALSE){
  plot(tree, drop = TRUE, tnex = 1.4,
       tp_args = list(fill = pal2[c(1, 3)], FUN = identity, width = 18), 
       ip_args = list(fill = pal2[c(2, 3)]))
}
@
\end{center}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Fitting a Tree}
\begin{center}
\vspace*{-0.5cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.74\linewidth}
<<plottree_box, fig=TRUE, echo=FALSE>>=
plot(as.constparty(tree), tnex = 1.4,
     tp_args = list(fill = pal2[c(1, 3)]), 
     ip_args = list(fill = pal2[c(2, 3)]))
@
\end{center}
\end{frame}

@@

  if(type == "exp"){
    #dev.off() 
    rmse <- cbind(simlist$av.rmse.exp.true.dt, simlist$av.rmse.exp.true.df, simlist$av.rmse.exp.true.g,
                  simlist$av.rmse.exp.obs.dt, simlist$av.rmse.exp.obs.df, simlist$av.rmse.exp.obs.g)
    colnames(rmse) <- c("dt.true", "df.true", "g.true", "dt.obs", "df.obs", "g.obs")
    ylim <- c(min(rmse), max(rmse))
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(x = simlist$x.axis, y = rmse[,"dt.true"], type = "l", col = pal["tree"], ylim = ylim,
         xlab = "", ylab = "", main = "disttree vs distforest vs gamlss", xaxt='n', yaxt='n', lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"dt.obs"], type = "l", lty = 2, col = pal["tree"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"df.true"], type = "l", col = pal["forest"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"df.obs"], type = "l", lty = 2, col = pal["forest"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"g.true"], type = "l", col = pal["gamlss"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"g.obs"], type = "l", lty = 2, col = pal["gamlss"], lwd = 2)
    #legend('left', c("disttree", "distforest", "gamlss"), col = c(pal["tree"], pal["forest"], pal["gamlss"]), lty = 1)
    mtext(text= "smooth", side = 1, line = 1, adj = 0)
    mtext(text= "steep", side = 1, line = 1, adj = 1)
    mtext(text= TeX('$\\kappa$'), side = 1, line = 1)
    mtext(text= "RMSE", side = 2, line = 1)
    mtext(text= "disttree", side = 2, col = pal["tree"], las = 1, line = 0.2, padj = 6.5)
    mtext(text= "distforest", side = 2, col = pal["forest"], las = 1, line = 0.2, padj = 13)
    mtext(text= "gamlss", side = 2, col = pal["gamlss"], las = 1, line = 0.3, padj = 16)

@@

    #add arrows below
    arrows(0.45,0.085,0.3,0.085,lwd=1) ## add arrow
    arrows(0.6,0.085,0.75,0.085,lwd=1) ## add arrow

@@

  if(type == "par"){
    rmse <- cbind(simlist$av.rmse.mu.dt, simlist$av.rmse.mu.df, simlist$av.rmse.mu.g, 
                  simlist$av.rmse.sigma.dt, simlist$av.rmse.sigma.df, simlist$av.rmse.sigma.g)
    colnames(rmse) <- c("mu.dt", "mu.df", "mu.g", "sigma.dt", "sigma.df", "sigma.g")
    ylim <- c(min(rmse), max(rmse))
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(x = simlist$x.axis, y = rmse[,"mu.dt"], type = "l", col = pal["tree"], ylim = ylim,
         xlab = "", ylab = "", main = "disttree vs distforest vs gamlss", xaxt='n', yaxt='n', lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"sigma.dt"], type = "l", lty = 2, col = pal["tree"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"mu.df"], type = "l", col = pal["forest"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"sigma.df"], type = "l", lty = 2, col = pal["forest"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"mu.g"], type = "l", col = pal["gamlss"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"sigma.g"], type = "l", lty = 2, col = pal["gamlss"], lwd = 2)
    #legend('topleft', c("disttree", "distforest", "gamlss"), col = c(pal["tree"], pal["forest"],pal["gamlss"]), lty = 1)
    mtext(text= "smooth", side = 1, line = 1, adj = 0)
    mtext(text= "steep", side = 1, line = 1, adj = 1)
    mtext(text= TeX('$\\kappa$'), side = 1, line = 1)
    mtext(text= "RMSE", side = 2, line = 1)
    mtext(text= "disttree", side = 2, col = pal["tree"], las = 1, line = 0.2, padj = -4)
    mtext(text= "distforest", side = 2, col = pal["forest"], las = 1, line = 0.2, padj = 8)
    mtext(text= "gamlss", side = 2, col = pal["gamlss"], las = 1, line = 0.3, padj = 14)

@@

# plot loglikelihood
plot_ll <- function(simlist){
  ll <- cbind(simlist$av.loglik.dt, simlist$av.loglik.df, simlist$av.loglik.g)
  colnames(ll) <- c("dt", "df","g")
  ylim <- c(min(ll), max(ll))
  par(mar=c(5.1,4.1,4.1,3.1))
  plot(x = simlist$x.axis, y = ll[,"dt"], type = "l", col = pal["tree"], ylim = ylim,
       xlab = "", ylab = "", main = "disttree vs distforest vs gamlss", xaxt='n', yaxt='n', lwd = 2)
  lines(x = simlist$x.axis, y = ll[,"df"], type = "l", col = pal["forest"], lwd = 2)
  lines(x = simlist$x.axis, y = ll[,"g"], type = "l", col = pal["gamlss"], lwd = 2)
  #legend('topright', c("disttree", "distforest", "gamlss"), col = c(pal["tree"], pal["forest"],pal["gamlss"]), lty = 1)
  mtext(text= "smooth", side = 1, line = 1, adj = 0)
  mtext(text= "steep", side = 1, line = 1, adj = 1)
  mtext(text= TeX('$\\kappa$'), side = 1, line = 1)
  mtext(text= "Log-Likelihood", side = 2, line = 1)
  mtext(text= "disttree", side = 2, col = pal["tree"], las = 1, line = 0.2, padj = 12)
  mtext(text= "distforest", side = 2, col = pal["forest"], las = 1, line = 0.2, padj = -8.5)
  mtext(text= "gamlss", side = 2, col = pal["gamlss"], las = 1, line = 0.3, padj = -16)
